---
title: "Find Features Stuff"
author: "Harriet Mason"
date: "04/02/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, loadlibraries}
library(tidyverse)
library(cassowaryr)
library(tourr)
library(ferrn)
library(stringr) 
```


# Feature Vs Noise
## Simulate Data
```{r}
# set seed
set.seed(2020)

# get L-shape data
lshape <- features %>% 
  filter(feature=="l-shape")

# make tibble group a
feature_vs_noise_A <- tibble(x1 = 2.5* cassowaryr:::unitize(lshape$x),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(lshape$y),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("A",100)
                           )

# Make tibble group b
feature_vs_noise_B <- tibble(x1 = rnorm(100, 0.5, 1),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = rnorm(100, 0.5, 1),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("B",100)
                           )


# combine tibbles
feature_vs_noise <- bind_rows(feature_vs_noise_A, feature_vs_noise_B)

# final tibble and ggpairs check
GGally::ggpairs(feature_vs_noise, columns = c(1:5),
                ggplot2::aes(colour = group, alpha=0.5))

```

## Calculate Scagnostics
```{r}
# calculate scagnostics
scags_feature_vs_noise_A <- calc_scags_wide(feature_vs_noise_A[,1:5]) %>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="A_value")

scags_feature_vs_noise_B <- calc_scags_wide(feature_vs_noise_B[,1:5])%>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="B_value")

# Combine Data
scags_feature_vs_noise_dif <- scags_feature_vs_noise_A %>%
  left_join(scags_feature_vs_noise_B) %>%
  mutate(dif_value = abs(A_value - B_value))

```

## Try Projection Persuit
Top differences were in convex, skinny, monotonic, and outlying
```{r, PP Functions}
# Scag Projection Persuit index
scags_groups <- function(cl, scag){
  classes <- unique(cl)
  function(mat){
    #each matrix group
    group1 <- mat[which(cl==classes[1]),]
    group2 <- mat[which(cl==classes[2]),]
    
    indx <- list( #copy and pasting should be cleaned up
      "convex" = (sc_convex(x = group1[,1], y = group1[,2]) - sc_convex(x = group2[,1], y = group2[,2])),
      "monotonic" = (sc_monotonic(x = group1[,1], y = group1[,2]) - sc_monotonic(x = group2[,1], y = group2[,2])),
      "outlying" = (sc_outlying(x = group1[,1], y = group1[,2]) - sc_outlying(x = group2[,1], y = group2[,2])),
      "splines" = (sc_splines(x = group1[,1], y = group1[,2]) - sc_splines(x = group2[,1], y = group2[,2])),
      "skinny" = (sc_skinny(x = group1[,1], y = group1[,2]) - sc_skinny(x = group2[,1], y = group2[,2]))
      )
    
    #final value
    abs(indx[[scag]])
  }
}

# Save Tour Function 
save_tourr <- function(numeric_data, group_vector, scag, search, seed, alpha, filename, ...){
  srch <- list(
    "bet" = save_history(numeric_data, tour_path = guided_tour(scags_groups(group_vector, scag), 
                                                               d=2, 
                                                               search_f =  search_better,
                                                               alpha = alpha,
                                                               ...)),
    "bet_rand" = save_history(numeric_data, tour_path = guided_tour(scags_groups(group_vector, scag),
                                                                    d=2,
                                                                    search_f =  search_better_random,
                                                                    alpha = alpha,
                                                               ...))
  )
  set.seed(seed)
  h <- srch[[search]]
  saveRDS(h, file=filename)
}

# Tour grid Function
grid_tourr <- function(numeric_data, group_vector, scags, searches, seeds, alpha, ...){
  expand.grid(scags, searches, seeds, alpha, stringsAsFactors=FALSE) %>%
    mutate(filename = paste("features_data/tours/tour", Var1, Var2, Var4, Var3, ".rds", sep="_")) %>%
    group_by(Var1, Var2, Var4, Var3) %>%
    summarise(save_tourr(numeric_data, group_vector, scag = Var1,  search = Var2, seed = Var3, 
                         alpha = Var4, filename = filename))
}
```

```{r, Grid Values, FvN}
# First Set of Overnight Values (Convex & Skinny)
grid_tourr(numeric_data = feature_vs_noise[1:5], 
           group_vector = feature_vs_noise$group,
           scags = c("convex", "skinny"), 
           searches = c("bet", "bet_rand"), 
           seeds = seq(5), 
           alpha = c(0.05, 0.1, 0.5, 0.7, 1)) 

# Second Set of Overnight Values (Monotonic and Outlying)
grid_tourr(numeric_data = feature_vs_noise[1:5], 
           group_vector = feature_vs_noise$group,
           scags = c("outlying", "monotonic"), 
           searches = c("bet", "bet_rand"), 
           seeds = seq(5), 
           alpha = c(0.1, 0.5, 0.7, 1)) 
# (Convex and Skinny)
grid_tourr(numeric_data = feature_vs_noise[1:5], 
           group_vector = feature_vs_noise$group,
           scags = c("convex", "skinny"), 
           searches = c("bet", "bet_rand"), 
           seeds = seq(5), 
           alpha = c(0.45, 0.55, 0.6, 0.65, 0.75))

```


# Feature Vs Feature
## Simulate Data
```{r}
# set seed
set.seed(2021)

# get L-shape data
lshape <- features %>% 
  filter(feature=="l-shape")

nonlinear <- features %>% 
  filter(feature=="nonlinear2")

# make tibble group a
feature_vs_feature_A <- tibble(x1 = 2.5* cassowaryr:::unitize(lshape$x),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(lshape$y),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("A",100)
                           )

# Make tibble group b
feature_vs_feature_B <- tibble(x1 = 2.5* cassowaryr:::unitize(nonlinear$x),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(nonlinear$y),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("B",100)
                           )


# combine tibbles
feature_vs_feature <- bind_rows(feature_vs_feature_A, feature_vs_feature_B)

# final tibble and ggpairs check
GGally::ggpairs(feature_vs_feature, columns = c(1:5),
                ggplot2::aes(colour = group, alpha=0.5))

```

## Calculate Scagnostics
```{r}
# calculate scagnostics
scags_feature_vs_feature_A <- calc_scags_wide(feature_vs_feature_A[,1:5]) %>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="A_value")

scags_feature_vs_feature_B <- calc_scags_wide(feature_vs_feature_B[,1:5])%>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="B_value")

# Combine Data
scags_feature_vs_feature_dif <- scags_feature_vs_feature_A %>%
  left_join(scags_feature_vs_feature_B) %>%
  mutate(dif_value = abs(A_value - B_value))
```

## Try Projection Persuit
Top differences were in splines, monotonic, and outlying
```{r, Grid Search FvF}
# (Convex and Skinny)
grid_tourr(numeric_data = feature_vs_feature[1:5], 
           group_vector = feature_vs_feature$group,
           scags = c("splines", "monotonic", "outlying"), 
           searches = c("bet", "bet_rand"), 
           seeds = seq(5), 
           alpha = c(0.1, 0.45, 0.55, 0.5, 0.6, 0.65, 0.7, 0.75, 1))

```



# Multiple Features Vs Noise
## Simulate Data
```{r}
# set seed
set.seed(2022)

# get L-shape data
lshape <- features %>% 
  filter(feature=="l-shape")

nonlinear <- features %>% 
  filter(feature=="nonlinear2")

# make tibble group a
mfeatures_vs_noise_A <- tibble(x1 = 2.5* cassowaryr:::unitize(lshape$x),
                           x2 = 2.5* cassowaryr:::unitize(nonlinear$x),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(lshape$y),
                           x5 = 2.5* cassowaryr:::unitize(nonlinear$y),
                           group = rep("A",100)
                           )

# Make tibble group b
mfeatures_vs_noise_B <- tibble(x1 = rnorm(100, 0.5, 1),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = rnorm(100, 0.5, 1),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("B",100)
                           )


# combine tibbles
mfeatures_vs_noise <- bind_rows(mfeatures_vs_noise_A, mfeatures_vs_noise_B)

# final tibble and ggpairs check
GGally::ggpairs(mfeatures_vs_noise, columns = c(1:5),
                ggplot2::aes(colour = group, alpha=0.5))

```
## Calculate Scagnostics
```{r}
# calculate scagnostics
scags_mfeatures_vs_noise_A <- calc_scags_wide(mfeatures_vs_noise_A[,1:5]) %>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="A_value")

scags_mfeatures_vs_noise_B <- calc_scags_wide(mfeatures_vs_noise_B[,1:5])%>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="B_value")

# Combine Data
scags_mfeatures_vs_noise_dif <- scags_mfeatures_vs_noise_A %>%
  left_join(scags_mfeatures_vs_noise_B) %>%
  mutate(dif_value = abs(A_value - B_value))
```

## Try Projection Persuit
Top differences were:
1) x1 vs x2: splines and dcor
2) x1 vs x4: convex

# Animate the Tours
````{r, Save and Trace which dont quite work}
# Animate Tour Function
anim_tourr <- function(numeric_data, group_vector, saved_tour, filename){
  a <- animate_xy(numeric_data,
                  tour_path = planned_tour(saved_tour),
                  col = group_vector)
  #saveRDS(a, filename)
}

#example
filename <- readRDS("features_data/tours2/tour_splines_bet_0.7_2_.rds")
anim_tourr(feature_vs_feature[1:5], feature_vs_feature$group, filename)

```

```{r}
# Save tour test
save_tourr(feature_vs_noise[1:5], feature_vs_noise$group, "monotonic", "geo", seed=1, alpha=0.01, "test1.rds")


# tour won't save stuff
t1 <- save_history(flea[, 1:6], max = 3)
a <- animate_xy(flea[, 1:6], planned_tour(t1))

b <- animate_xy(feature_vs_noise[1:5],
                tour_path = guided_tour(scags_groups(feature_vs_noise$group, "convex")),
                col = feature_vs_noise$group)

c <- animate(feature_vs_noise[1:5],
                tour_path = planned_tour(goodvibes),
                col = feature_vs_noise$group)

d <- animate_xy(feature_vs_noise[1:5], max = 30,
             col = feature_vs_noise$group)


saveRDS(c, "badvibes.rds")

# Tour Trace Plot Function
trace_plot <- function(saved_anim, plot_name, filename){
  p <- saved_anim %>%
    explore_trace_interp() + 
    scale_color_continuous_botanical() + 
    ggtitle(paste("plotname"))
  ggsave(filename, p)
}

#example
badvibes <- readRDS("C:/Users/harri/OneDrive/2021_Summer_Research/badvibes.rds")
trace_plot(c, "Vibes Plot Check", "neutralvibes.jpg")


```
