---
title: "Find Features Stuff"
author: "Harriet Mason"
date: "04/02/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, loadlibraries}
library(tidyverse)
library(cassowaryr)
library(tourr)
library(ferrn)
library(stringr)  
```

# The feature vs noise data

We want to identify if scagnostics can be used in projection persuit to find interesting shapes in linear projections of a data set. In order to test this, we created a data set with two groups where one pair of variables has a "shape" in one of the groups, and all other variables are normally distributed. Here the scagnsotics will have to find the L-shape in group A in the x1 and x4 pairwise plot that can be seen in the scatter plot matrix (SPLOM) below. If the scagnsotics can be used in projection persuit, they should be able to identify which group and variable combination shows the abonormal shape.

```{r simdata, fig.cap= "A scatter plot matrix of the simulated 'L-shape vs noise' data. The shape can be seen in the x1 vs x4 pairwise plot in group A."}
# set seed
set.seed(2020)

# get L-shape data
lshape <- features %>% 
  filter(feature=="l-shape")

# make tibble group a
feature_vs_noise_A <- tibble(x1 = 2.5* cassowaryr:::unitize(lshape$x),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(lshape$y),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("A",100)
                           )

# Make tibble group b
feature_vs_noise_B <- tibble(x1 = rnorm(100, 0.5, 1),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = rnorm(100, 0.5, 1),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("B",100)
                           )


# combine tibbles
feature_vs_noise <- bind_rows(feature_vs_noise_A, feature_vs_noise_B)

# final tibble and ggpairs check
p1 <- GGally::ggpairs(feature_vs_noise, columns = c(1:5),
                ggplot2::aes(colour = group, alpha=0.5))+ 
  theme_minimal()
  

p1
```

Any scagnsotic that can identify this shape in projection pursuit should also be able to pick the shape from a collection pairwise plots. This is how we are going to select the projection pursuit index. Here we calculate each scagnostic, for each possible pairwise plot, separately for each group. We then calculate the absolute difference between these values for each group. The scagnostics that have their maximum group difference on the x1 and x4 pair can "see" the shape. The table below has the difference between the groups and the associated variable pair for each scagnostic. The scagnostics convex and skinny are the best choice for projection pursuit and we will continue with those measures.

```{r, eval=FALSE scagcals, fig.cap= "The top pairs of variables. Convex and skinny do the best job of seeing the L-shape in group A. }
# calculate scagnostics
scags_feature_vs_noise_A <- calc_scags_wide(feature_vs_noise_A[,1:5]) %>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="A_value")

scags_feature_vs_noise_B <- calc_scags_wide(feature_vs_noise_B[,1:5])%>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="B_value")

# Combine Data
scags_feature_vs_noise_dif <- scags_feature_vs_noise_A %>%
  left_join(scags_feature_vs_noise_B) %>%
  mutate(dif_value = abs(A_value - B_value)) 

scag_df <- scags_feature_vs_noise_dif %>%
  select(-c(A_value, B_value)) %>%
  pivot_wider(id_cols = c(Var1, Var2), names_from=scag, values_from=dif_value)
  
top_scags(scag_df) %>%
  arrange(desc(value))
  
```

## 2) The Functions (Using the Projection Persuit Index)
### a) projections persuit
### b) The functions I ended up using
## 3) Tuning the Projection Persuit Indexes
## 4) Trace plots and Results
## 4) Improvements / Which scagnostics can be used in projection persuit + why
- we have worked backwards here to show that the scagnsotics can be used to identify shapes and tinker the optimiser. When using scagnostics in projection persuit you would know what kind of shape you are looking for, and do a projection persuit to see if it exists anywhere in the data (While the "correct: projection would not usually be known to us, since we are trying to see if the measures work at all, we will be working backwards from the final results)
- clumpy2 is too computationally expensive
- not scaled globally 0 to 1
- scagnstics do not actually scale 0 to 1, therefor the scagnostic with the biggest difference may not be convex, but sparse. cannot tell because of scaling issue.
- more variables
- different shapes


##################################### SANDBOX WORK #########################################
## Try Projection Persuit
Top differences were in convex, skinny, monotonic, and outlying

```{r, eval=FALSE, Projection Persuit Functions}
# Record
options("tourr.verbose" = TRUE)

# Scag Projection Persuit index
scags_groups <- function(cl, scag){
  classes <- unique(cl)
  function(mat){
    #each matrix group
    group1 <- mat[which(cl==classes[1]),]
    group2 <- mat[which(cl==classes[2]),]
    
    indx <- list( #copy and pasting should be cleaned up
      "convex" = (sc_convex(x = group1[,1], y = group1[,2]) - sc_convex(x = group2[,1], y = group2[,2])),
      "monotonic" = (sc_monotonic(x = group1[,1], y = group1[,2]) - sc_monotonic(x = group2[,1], y = group2[,2])),
      "outlying" = (sc_outlying(x = group1[,1], y = group1[,2]) - sc_outlying(x = group2[,1], y = group2[,2])),
      "splines" = (sc_splines(x = group1[,1], y = group1[,2]) - sc_splines(x = group2[,1], y = group2[,2])),
      "skinny" = (sc_skinny(x = group1[,1], y = group1[,2]) - sc_skinny(x = group2[,1], y = group2[,2]))
      )
    
    #final value
    abs(indx[[scag]])
  }
}

# Save Tour Function 
save_tourr <- function(numeric_data, group_vector, scag, search, seed, alpha, max.tries, filename, ...){
  #set seed
  set.seed(seed)
  
  srch <- list(
    "bet" = save_history(numeric_data, tour_path = guided_tour(scags_groups(group_vector, scag), 
                                                               d=2, 
                                                               search_f =  search_better,
                                                               alpha = alpha,
                                                               max.tries = max.tries,
                                                               ...)),
    "bet_rand" = save_history(numeric_data, tour_path = guided_tour(scags_groups(group_vector, scag),
                                                                    d=2,
                                                                    search_f =  search_better_random,
                                                                    alpha = alpha,
                                                                    max.tries = max.tries,
                                                                    ...))
  )
  h <- srch[[search]]
  saveRDS(h, file=filename)
}

# Tour grid Function
grid_tourr <- function(numeric_data, group_vector, scags, searches, seeds, alpha, max.tries, filestart, ...){
  #make expanded tibble so that we can search a grid of values
  expand.grid(scags, searches, seeds, alpha, max.tries, stringsAsFactors=FALSE) %>%
    # add file names for saving
    mutate(filename = paste(filestart, Var1, Var2, Var4, Var5, Var3, ".rds", sep="_")) %>%
    # group by and summarise to run save on every value
    group_by(Var1, Var2, Var4, Var5, Var3) %>%
    summarise(save_tourr(numeric_data, group_vector, scag = Var1,  search = Var2, seed = Var3, 
                         alpha = Var4, max.tries=Var5, filename = filename))
}


# Animate Tour Function
anim_tourr <- function(numeric_data, group_vector, scag, search, seed, alpha, max.tries, filestart){
  #set seed
  set.seed(seed)
  
  # perform search by calling function from list
  srch <- list(
    "bet" =  animate_xy(numeric_data,
                  tour_path = guided_tour(scags_groups(group_vector, scag),
                                          search_f = search_better,
                                          alpha = alpha,
                                          max.tries=max.tries),
                  col = group_vector),
    "bet_rand" = animate_xy(numeric_data,
                  tour_path = guided_tour(scags_groups(group_vector, scag),
                                          search_f = search_better_random,
                                          alpha = alpha,
                                          max.tries=max.tries),
                  col = group_vector)
    )
  
  #save object and file name
  a <- srch[[search]]
  filename <- paste(filestart, scag, search, alpha, max.tries, seed, ".rds", sep="_")
  
  # save r object
  saveRDS(a, filename)
}

# Trace Plot Function
trace_plot <- function(saved_anim, plot_name, filename){
  p <- saved_anim %>%
    explore_trace_interp() + 
    scale_color_continuous_botanical() + 
    ggtitle(paste("plotname"))
  ggsave(filename, p)
}

# Polish function
save_polish <- function(numeric_data, group_vector, scag, seed = 1, 
                        polish_max_tries = 30, saved_tour_name, filename){
  
  # get final projection of saved tour
  attr(saved_tour, "class") <- NULL
  best_proj <- saved_tour[, , dim(saved_tour)[3]]
  
  # set seed
  set.seed(seed)
  
  # run search
  h <- save_history(numeric_data, 
                    tour_path = guided_tour(scags_groups(group_vector, scag),
                                            search_f = search_polish,
                                            polish_max_tries = polish_max_tries,
                                            cur_index = 0),
                    start = best_proj)
  saveRDS(h, file=filename)
  
}



```


````{r, eval=FALSE, Function Tests}
#save_tourr test
save_tourr(feature_vs_noise[1:5], feature_vs_noise$group, "convex", "bet", seed=1, alpha=1, max.tries = 25,
           filename = "features_data/tours/tour_convex_bet_1_25_1_.rds")

# grid search test
grid_tourr(numeric_data = feature_vs_noise[1:5], 
           group_vector = feature_vs_noise$group,
           scags = c("convex"), 
           searches = c("bet"), 
           seeds = 1, 
           alpha = 0.4,
           max.tries = 100,
           filestart = "features_data/tours/tour")


# anim_tourr test
anim_tourr(feature_vs_noise[1:5], feature_vs_noise$group, "convex", "bet", seed=1, 
           alpha=0.5, filestart = 'features_data/anims/anim')


# Test POLISH
save_polish(numeric_data = feature_vs_noise[1:5], 
            group_vector = feature_vs_noise$group, 
            scag = "convex", 
            #seed = 1, 
            #polish_max_tries, 
            saved_tour = readRDS("features_data/tours1/tour_skinny_bet_7_500_5_.rds"), 
            filename = "features_data/tours1/tour_skinny_bet_7_500_5_(polish).rds")
  # watch
saved_tour <- NULL
saved_tour <- readRDS("features_data/tours1/tour_skinny_bet_7_500_5_.rds")
animate_xy(feature_vs_noise[1:5], 
        tour_path = planned_tour(saved_tour),
        col = feature_vs_noise$group)

saved_tour <- NULL
saved_tour <- readRDS("features_data/tours1/tour_skinny_bet_7_500_5_(polish).rds")
animate_xy(feature_vs_noise[1:5], 
        tour_path = planned_tour(saved_tour),
        col = feature_vs_noise$group)

# Trace plot test

```

```{r, eval=FALSE, Grid Values, FvN}
filestart <- "features_data/tours/tour"
# First Set of Overnight Values (Convex & Skinny)
grid_tourr(numeric_data = feature_vs_noise[1:5], 
           group_vector = feature_vs_noise$group,
           scags = c("convex", "skinny"), 
           searches = c("bet", "bet_rand"), 
           seeds = seq(5), 
           alpha = c(0.05, 0.1, 0.5, 0.7, 1),
           max.tries = 25,
           filestart = filestart) 
# Second Set of Overnight Values (Monotonic and Outlying)
grid_tourr(numeric_data = feature_vs_noise[1:5], 
           group_vector = feature_vs_noise$group,
           scags = c("outlying", "monotonic"), 
           searches = c("bet", "bet_rand"), 
           seeds = seq(5), 
           alpha = c(0.1, 0.5, 0.7, 1),
           max.tries = 25,
           filestart = filestart) 
# (Convex and Skinny)
grid_tourr(numeric_data = feature_vs_noise[1:5], 
           group_vector = feature_vs_noise$group,
           scags = c("convex", "skinny"), 
           searches = c("bet", "bet_rand"), 
           seeds = seq(5), 
           alpha = c(0.45, 0.55, 0.6, 0.65, 0.75),
           max.tries = 25,
           filestart = filestart)

# Adjust Max Size Grid Search 
grid_tourr(numeric_data = feature_vs_noise[1:5], 
           group_vector = feature_vs_noise$group,
           scags = c("convex", "skinny"), #convex and skinny were the best of the scagnsotics
           searches = "bet",  #better was the better search lol
           seeds = seq(10), #increase number of seeds to increase certainty
           alpha = c(0.1, 0.3, 0.5, 0,7), #can have a lower alpha with higher max.tries
           max.tries = c(50, 100, 200, 300, 500), #test max tries
           filestart = "features_data/tours1/tour") #new folder so not with others

```

```{r, eval=FALSE, Working Examples for animate_xy}
# GEODESIC WORK
set.seed(2)
m <- animate_xy(feature_vs_noise[1:5],
                tour_path = guided_tour(scags_groups(feature_vs_noise$group, "convex")),
                delta=0.05,
                col = feature_vs_noise$group)
 # BETTER WORK
set.seed(7)
m <- animate_xy(feature_vs_noise[1:5],
                tour_path = guided_tour(scags_groups(feature_vs_noise$group, "convex"),
                                        search_f = search_better,
                                        alpha=0.75),
                col = feature_vs_noise$group)

# Anim_tour version of better work that does not produce the same results
anim_tourr(feature_vs_noise[1:5], feature_vs_noise$group, "convex", "bet", seed=7, 
           alpha=0.75, filestart = 'features_data/anims/anim')
```

# Feature Vs Feature
## Simulate Data
```{r, eval=FALSE}
# set seed
set.seed(2021)

# get L-shape data
lshape <- features %>% 
  filter(feature=="l-shape")

nonlinear <- features %>% 
  filter(feature=="nonlinear2")

# make tibble group a
feature_vs_feature_A <- tibble(x1 = 2.5* cassowaryr:::unitize(lshape$x),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(lshape$y),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("A",100)
                           )

# Make tibble group b
feature_vs_feature_B <- tibble(x1 = 2.5* cassowaryr:::unitize(nonlinear$x),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(nonlinear$y),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("B",100)
                           )


# combine tibbles
feature_vs_feature <- bind_rows(feature_vs_feature_A, feature_vs_feature_B)

# final tibble and ggpairs check
GGally::ggpairs(feature_vs_feature, columns = c(1:5),
                ggplot2::aes(colour = group, alpha=0.5))

```

## Calculate Scagnostics
```{r, eval=FALSE}
# calculate scagnostics
scags_feature_vs_feature_A <- calc_scags_wide(feature_vs_feature_A[,1:5]) %>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="A_value")

scags_feature_vs_feature_B <- calc_scags_wide(feature_vs_feature_B[,1:5])%>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="B_value")

# Combine Data
scags_feature_vs_feature_dif <- scags_feature_vs_feature_A %>%
  left_join(scags_feature_vs_feature_B) %>%
  mutate(dif_value = abs(A_value - B_value))
```

## Try Projection Persuit
Top differences were in splines, monotonic, and outlying
```{r, eval=FALSE, Grid Search FvF}
# (Convex and Skinny)
grid_tourr(numeric_data = feature_vs_feature[1:5], 
           group_vector = feature_vs_feature$group,
           scags = c("splines", "monotonic", "outlying"), 
           searches = c("bet", "bet_rand"), 
           seeds = seq(5), 
           alpha = c(0.1, 0.45, 0.55, 0.5, 0.6, 0.65, 0.7, 0.75, 1))

```



# Multiple Features Vs Noise
## Simulate Data
```{r, eval=FALSE}
# set seed
set.seed(2022)

# get L-shape data
lshape <- features %>% 
  filter(feature=="l-shape")

nonlinear <- features %>% 
  filter(feature=="nonlinear2")

# make tibble group a
mfeatures_vs_noise_A <- tibble(x1 = 2.5* cassowaryr:::unitize(lshape$x),
                           x2 = 2.5* cassowaryr:::unitize(nonlinear$x),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(lshape$y),
                           x5 = 2.5* cassowaryr:::unitize(nonlinear$y),
                           group = rep("A",100)
                           )

# Make tibble group b
mfeatures_vs_noise_B <- tibble(x1 = rnorm(100, 0.5, 1),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = rnorm(100, 0.5, 1),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("B",100)
                           )


# combine tibbles
mfeatures_vs_noise <- bind_rows(mfeatures_vs_noise_A, mfeatures_vs_noise_B)

# final tibble and ggpairs check
GGally::ggpairs(mfeatures_vs_noise, columns = c(1:5),
                ggplot2::aes(colour = group, alpha=0.5))

```
## Calculate Scagnostics
```{r, eval=FALSE}
# calculate scagnostics
scags_mfeatures_vs_noise_A <- calc_scags_wide(mfeatures_vs_noise_A[,1:5]) %>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="A_value")

scags_mfeatures_vs_noise_B <- calc_scags_wide(mfeatures_vs_noise_B[,1:5])%>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="B_value")

# Combine Data
scags_mfeatures_vs_noise_dif <- scags_mfeatures_vs_noise_A %>%
  left_join(scags_mfeatures_vs_noise_B) %>%
  mutate(dif_value = abs(A_value - B_value))
```

## Try Projection Persuit
Top differences were:
1) x1 vs x2: splines and dcor
2) x1 vs x4: convex

# Watching the Tours

```{r, eval=FALSE}
# watch tour
# format: tour_scag_search_alpha_maxtries_seed_.rds
saved_tour <- NULL
saved_tour <- readRDS("features_data/tours1/tour_skinny_bet_7_500_5_.rds")
animate_xy(feature_vs_noise[1:5], 
        tour_path = planned_tour(saved_tour),
        col = feature_vs_noise$group)

```

# Make trace plot function from Save History
```{r, eval=FALSE}
get_trace <- function(){
  
}

tPath <- readRDS("features_data/tours1/tour_skinny_bet_7_500_5_.rds")
tFullPath <- as.list(tourr::interpolate(tPath))

```

# Get last frame from animation
```{r, eval=FALSE}
last_frame <- function(numeric_data, group_data, anim_data){
  # make data into matrix
  mat_data <- as.matrix(numeric_data)
  mat_basis <- matrix(unlist(get_best(anim_data)$basis), ncol=2, nrow=5)
  
  # transform data to basis
  plot_data <- as_tibble(mat_data %*% mat_basis) %>%
    rename(x = V1, y = V2)
  plot_data$group <- group_data
  
  # get basis data for segment plots
  basis_data <- as_tibble(matrix(unlist(get_best(anim_data)$basis), ncol=2, nrow=5)) %>%
    rename(xend = V1, yend = V2) %>%
    mutate(x=0, y=0, xend=5*xend, yend=5*yend)
  basis_data$variable = c("x1", "x2", "x3", "x4", "x5")
  
  #plot ggplot
  ggplot(plot_data, aes(x,y)) + 
    geom_point(aes(colour=group)) +
    theme_classic() +
    geom_segment(data=basis_data, aes(x=x, y=y, xend=xend, yend=yend)) +
    geom_text(data=basis_data, aes(x=xend, y=yend, label=variable)) +
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank())

}


saved_anim <- anim_tourr(feature_vs_noise[1:5], feature_vs_noise$group, "skinny", "bet", 7, 0.7, 500, "anim")
saved_tour <- NULL
saved_anim <- readRDS("anim_skinny_bet_0.7_500_7_.rds")
last_frame(feature_vs_noise[1:5], feature_vs_noise$group, check)

numeric_data <- feature_vs_noise[1:5]
group_data <- feature_vs_noise$group
mat_data <- as.matrix(numeric_data)
  mat_basis <- matrix(unlist(check$basis), ncol=2, nrow=5) #matrix(unlist(get_best(check)$basis), ncol=2, nrow=5)
  
  # transform data to basis
  plot_data <- as_tibble(mat_data %*% mat_basis) %>%
    rename(x = V1, y = V2)
  plot_data$group <- group_data
  
  # get basis data for segment plots
  basis_data <- as_tibble(matrix(unlist(check$basis), ncol=2, nrow=5)) %>%
    rename(xend = V1, yend = V2) %>%
    mutate(x=0, y=0, xend=5*xend, yend=5*yend)
  basis_data$variable = c("x1", "x2", "x3", "x4", "x5")
  
  #plot ggplot
  ggplot(plot_data, aes(x,y)) + 
    geom_point(aes(colour=group)) +
    theme_classic() +
    geom_segment(data=basis_data, aes(x=x, y=y, xend=xend, yend=yend)) +
    geom_text(data=basis_data, aes(x=xend, y=yend, label=variable)) +
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank())


```




