---
title: "Try to Identify Numbat in Noise"
author: "Harriet Mason"
date: "07/12/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load package and data
```{r}
# cassowaryr
library(cassowaryr)

#data wrangling
library(tidyverse)

# pairwise plots
library(GGally)
library(plotly)

#data partition 
library(caret)

# add in tourr
library(tourr)

# load data
numbats_raw <- read.csv("numbat.csv")
numbats_raw[,1:10] <- rescale(numbats_raw[,1:10])
```

# Whole Data SPLOM
```{r}
ggpairs(numbats_raw, columns=1:11, ggplot2::aes(colour=group, alpha=0.1))
```
- difficult to see anything because the blue group is over the top. Maybe with a smaller sample or something I could see it. Can kind of tell its there though.

## Calculate between group differences in all scagnostics
First calculate the scagnostics on the entire dataset (save data because this takes forever).

## No outlier removal
As the functions are when I call them with sc_ (which I did for some)
```{r, eval=FALSE}
# make group data frames
bignumbatsA <- filter(numbats_raw, group=="A") %>% select(-group)
bignumbatsB <- filter(numbats_raw, group=="B") %>% select(-group)

#calc_scags on both groups
bignumbatsA_scags_out <- calc_scags_wide(bignumbatsA, global_rng=c(0,1), out.rm=FALSE)  
bignumbatsB_scags_out <- calc_scags_wide(bignumbatsB, global_rng=c(0,1), out.rm=FALSE)

saveRDS(bignumbatsA_scags_out, "bigrawscagsA_out.rds")
saveRDS(bignumbatsB_scags_out, "bigrawscagsB_out.rds")
```

```{r}
#combine into a single data frame
bignumbatsA_scags_outrm_long <- bignumbatsA_scag_outrms %>% 
  pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "A_value")

bignumbatsB_scags_outrm_long <- bignumbatsB_scags_outrm %>% 
  pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "B_value")

#relative dist
bignumbats_scags_long_dist1 <- bignumbatsA_scags_outrm_long %>%
  left_join(bignumbatsB_scags_outrm_long, by=c("Var1", "Var2", "scag")) %>%
  mutate(group_diff = (A_value - B_value)/(A_value + B_value))

#make wide data
bignumbats_scags_wide_1 <- bignumbats_scags_long_dist1 %>%
  select(c(Var1, Var2, scag, group_diff)) %>%
  pivot_wider(id_cols=c(Var1, Var2), names_from=scag, values_from = group_diff)

#save data
saveRDS(bignumbats_scags_wide, "wholescags_int_out.rds")

#abs dist
bignumbats_scags_long_dist2 <- bignumbatsA_scags_outrm_long %>%
  left_join(bignumbatsB_scags_outrm_long, by=c("Var1", "Var2", "scag")) %>%
  mutate(group_diff = abs(A_value - B_value))


         #make wide data
bignumbats_scags_wide_2 <- bignumbats_scags_long_dist2 %>%
  select(c(Var1, Var2, scag, group_diff)) %>%
  pivot_wider(id_cols=c(Var1, Var2), names_from=scag, values_from = group_diff)

#save data
saveRDS(bignumbats_scags_wide_2, "wholescags_abs_out.rds")

```

## Outlier Removal
Even though removing outliers is more computationally heavy, if there is enough of a gap might be worth doing calc_scags instead of sc_

```{r, eval=FALSE}
# make group data frames
bignumbatsA <- filter(numbats_raw, group=="A") %>% select(-group)
bignumbatsB <- filter(numbats_raw, group=="B") %>% select(-group)

#calc_scags on both groups
bignumbatsA_scags_outrm <- calc_scags_wide(bignumbatsA, global_rng=c(0,1))  
bignumbatsB_scags_outrm <- calc_scags_wide(bignumbatsB, global_rng=c(0,1))


saveRDS(bignumbatsA_scags_outrm, "bigrawscagsA_outrm.rds")
saveRDS(bignumbatsB_scags_outrm, "bigrawscagsB_outrm.rds")
```

```{r}
#combine into a single data frame
bignumbatsA_scags_outrm_long <- bignumbatsA_scags_outrm %>% 
  pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "A_value")

bignumbatsB_scags_outrm_long <- bignumbatsB_scags_outrm %>% 
  pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "B_value")

#RELATIVE DISTANCE
bignumbats_scags_long_dist3 <- bignumbatsA_scags_outrm_long %>%
  left_join(bignumbatsB_scags_outrm_long, by=c("Var1", "Var2", "scag")) %>%
  mutate(group_diff = (A_value - B_value)/(A_value + B_value))
  # wide data
bignumbats_scags_wide_3 <- bignumbats_scags_long_dist3 %>%
  select(c(Var1, Var2, scag, group_diff)) %>%
  pivot_wider(id_cols=c(Var1, Var2), names_from=scag, values_from = group_diff)

#save data
saveRDS(bignumbats_scags_wide_3, "wholescags_int_outrm.rds")

#abs dist
bignumbats_scags_long_dist4 <- bignumbatsA_scags_outrm_long %>%
  left_join(bignumbatsB_scags_outrm_long, by=c("Var1", "Var2", "scag")) %>%
  mutate(group_diff = abs(A_value - B_value))
  #make wide data
bignumbats_scags_wide_4 <- bignumbats_scags_long_dist4 %>%
  select(c(Var1, Var2, scag, group_diff)) %>%
  pivot_wider(id_cols=c(Var1, Var2), names_from=scag, values_from = group_diff)

#save data
saveRDS(bignumbats_scags_wide_4, "wholescags_abs_outrm.rds")

```

## Try Relative Distance with a Threshold
```{r}
#load data
bignumbatsA_scags_outrm <- readRDS("bigrawscagsA_outrm.rds")
bignumbatsB_scags_outrm <- readRDS("bigrawscagsB_outrm.rds")

# set upper range of "nothing" group functions
bignumbatsA_scags_outrm_long <- bignumbatsA_scags_outrm %>% 
  pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "A_value")

bignumbatsB_scags_outrm_long <- bignumbatsB_scags_outrm %>% 
  pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "B_value")

# Threshold function
scag_threshold <- function(scag, value){
  #set thresholds
  thresholds <- list(
    "clumpy2" = 0.5,
    "convex" = 0.59, # numbat identified by a small value (actually invert)
    "dcor" = 0.06, 
    "monotonic" = 0.06,
    "outlying"= 0.25, # numbat is identified by a small value
    "skewed" = 0.65, # numbat is identified by a small value
    "skinny" = 0.5, 
    "sparse" = 0.016, # numbat is identified by a small value
    "splines" = 0.02,
    "striated2" = 0.1,
    "stringy" = 0.75
  )
  #inverse values
  
  #return value
  if (scag=="convex") value #ifelse(value > thresholds[scag], 0, value)
  else ifelse(value < thresholds[scag], 0, value)
  
}

# Compute Data
bignumbats_scags_long_dist <- bignumbatsA_scags_outrm_long %>%
  left_join(bignumbatsB_scags_outrm_long, by=c("Var1", "Var2", "scag")) %>%
  # Replace with threshold values
  mutate(A_value = scag_threshold(scag, A_value),
         B_value = scag_threshold(scag, B_value),
         # calc group difference (note if both are 0)
         group_diff = ifelse((A_value + B_value)==0, 0, (A_value - B_value)/(A_value + B_value))
  )


bignumbats_scags_wide <- bignumbats_scags_long_dist%>%
  select(c(Var1, Var2, scag, group_diff)) %>%
  pivot_wider(id_cols=c(Var1, Var2), names_from=scag, values_from = group_diff)

#visualise


```

# Which Combinations of Scagnsotics could be used

```{r}

#look at SPLOM of scagnostics
splom_data <- bignumbats_scags_wide %>%
  mutate(lab = paste0(Var1, " , ", Var2)) %>%
  ungroup() %>%
  mutate(numbat=ifelse(lab == "x7 , x4", TRUE, FALSE)) %>%
  select(-c(Var1, Var2))

#interactive ggpairs
p <- ggpairs(splom_data, columns=1:11, ggplot2::aes(label=lab)) +
  theme_minimal()
ggplotly(p) 
```

## Absolute dist
- Aren't really any scagnostics where it sits away from the general "trend" of the data. Mostly would need to use individual or combinations of scagnsotics.
# Scaled Dist
- Monotonic and dcor kind of do?

# Which individual scagnostics should be used?

```{r}
# top pair for each scag
top_scags(bignumbats_scags_wide_3)
```

## Absolute group
Scag Placings:
- Numbat 1st: clumpy, Dcor, monotonic, outlying splines
- Numbat 2nd: Convex, Skinny, Sparse
- Numbat top 5: Skewed
- Numbat Nowhere: Euclid, Striated2, Stringy

Does it change with sample/sample size? Yes

Does it change with outlier removal? Yes
- x7 and x4 now top clumpy 2 difference???

Clumpy2 Specific
- identifies variable 7 as something interesting but not the combination. 
- Invites an investigation that looks at only one variable
- Alternatively could compare groups to the average across all variables? 
- Still maintains issue that one variable highly influences this scagnsotic, also 

Euclid
- Doesn't appear at all on euclidean distance
- also makes me think the clumpy value's sharp drop to 0 influences things like euclidean distance too much

## Scaled group differnce
Now only Dcor and Splines have x2/x7 on top


### Redo SPLOM and top-scags checks
```{r}
bignumbats_scags_wide_out <- readRDS("wholesetscags_outliers.rds")

#look at SPLOM of scagnostics
splom_data_out <- bignumbats_scags_wide_out %>%
  mutate(lab = paste0(Var1, " , ", Var2)) %>%
  ungroup() %>%
  mutate(numbat=ifelse(lab == "x7 , x4", TRUE, FALSE)) %>%
  select(-c(Var1, Var2))

#interactive ggpairs
p <- ggpairs(splom_data_out, columns=1:11, ggplot2::aes(label=lab)) +
  theme_minimal()
ggplotly(p) 
```

```{r}
# top pair for each scag
top_scags(bignumbats_scags_wide_out)
```
- after keeping outliers, it is still only dcor and splines that have numbat view on top
- that being said, the splom shows that it is harder to identify splines
- maybe should try using dcor instead of splines, it is more consistent?
- clumpy2 is also behaving really weirdly. I think it is trying to divide by for several of them, rest just return a 1 tho for some reason.
- skewed, sparse and convex also seem to be able to kind of see it, so maybe i should also try those.

## How much noise do we have on a 500 observation sample

```{r, eval=FALSE}
#subset variables
numbats_small <- numbats_raw %>%
  select(x1,x2,x4,x7, group)

# Set data
scagnoise_calc <- function(n, out.rm, intdist){
  scag_data <- NULL
  for(i in seq(n)){
    # load data
    set.seed(i)
    smallset <- createDataPartition(numbats_small$group, p=0.25)$Resample1
    numbatset <- numbats_small[smallset,]
    
    # make group data frames
    numbatsA <- filter(numbatset, group=="A") %>% select(-group)
    numbatsB <- filter(numbatset, group=="B") %>% select(-group)
    
    #calc_scags on both groups
    numbatsA_scags <- calc_scags_wide(numbatsA, global_rng=c(0,1), out.rm = out.rm)  
    numbatsB_scags <- calc_scags_wide(numbatsB, global_rng=c(0,1), out.rm = out.rm)
    
    #combine into a single data frame
    numbatsA_scags_long <- numbatsA_scags %>% 
      pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "A_value")
    
    numbatsB_scags_long <- numbatsB_scags %>% 
      pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "B_value")
    
    if(intdist) {
      numbats_scags_long <- numbatsA_scags_long %>%
        left_join(numbatsB_scags_long, by=c("Var1", "Var2", "scag")) %>%
        mutate(group_diff = (A_value - B_value)/(A_value + B_value))
    }
    if(!intdist){
      numbats_scags_long <- numbatsA_scags_long %>%
        left_join(numbatsB_scags_long, by=c("Var1", "Var2", "scag")) %>%
        mutate(group_diff = abs(A_value - B_value))
    }
    #make wide data
    numbats_scags_wide <- numbats_scags_long %>%
      select(c(Var1, Var2, scag, group_diff)) %>%
      pivot_wider(id_cols=c(Var1, Var2), names_from=scag, values_from = group_diff)
    
    # check 
    scag_data <- bind_rows(scag_data, numbats_scags_wide)
  }
  scag_data
}

scag_noise_1 <- scagnoise_calc(20, FALSE, TRUE)
scag_noise_2 <- scagnoise_calc(20, FALSE, FALSE)
scag_noise_3 <- scagnoise_calc(20, out.rm= TRUE, intdist = TRUE)
scag_noise_4 <- scagnoise_calc(20, out.rm = TRUE, intdist = FALSE)

saveRDS(scag_noise_1, "scagnoise_int_out.rds")
saveRDS(scag_noise_2, "scagnoise_abs_out.rds")
saveRDS(scag_noise_3, "scagnoise_int_outrm.rds")
saveRDS(scag_noise_4, "scagnoise_abs_outrm.rds")
```


```{r}
# Plot Function
generateplot <- function(scag_data, bignumbats_scags_wide, title){
  #tidy true values
  smallsetplots <- c("x2 , x1", "x4 , x1", "x4 , x2", 
                     "x7 , x1", "x7 , x2", "x7 , x4")
  scag_true <- bignumbats_scags_wide %>% 
    mutate(lab = paste0(Var1, " , ", Var2)) %>%
    select(-c(Var1, Var2)) %>%
    filter(lab %in% smallsetplots)%>%
    pivot_longer(cols=-lab, names_to = "scag", values_to = "value")
  
  #tidy data
  scag_noise <- scag_data %>% 
    mutate(lab = paste0(Var1, " , ", Var2)) %>%
    select(-c(Var1, Var2)) %>%
    pivot_longer(cols=-lab, names_to = "scag", values_to = "value")
  
  #plot
  ggplot(scag_noise, aes(x=lab, y=value, colour=lab)) +
    geom_violin()+
    geom_point(alpha=0.5)+
    facet_wrap(~scag, scales = "free") +
    theme_minimal() +
    geom_point(data = scag_true, colour="black") +
    ggtitle(title)
}

#Load Data
scag_noise_1 <- readRDS("scagnoise_int_out.rds")
bignumbats_scags_wide_1 <- readRDS("wholescags_int_out.rds")
scag_noise_2 <- readRDS("scagnoise_abs_out.rds")
bignumbats_scags_wide_2 <- readRDS("wholescags_abs_out.rds")
scag_noise_3 <- readRDS("scagnoise_int_outrm.rds")
bignumbats_scags_wide_3 <- readRDS("wholescags_int_outrm.rds")
scag_noise_4 <- readRDS("scagnoise_abs_outrm.rds")
bignumbats_scags_wide_4 <- readRDS("wholescags_abs_outrm.rds")

# Generate Plots
p1 <- generateplot(scag_noise_1, bignumbats_scags_wide_1, "Relative Distance and No Outlier Removal")
p2 <- generateplot(scag_noise_2, bignumbats_scags_wide_2, "Absolute Distance and No Outlier Removal")
p3 <- generateplot(scag_noise_3, bignumbats_scags_wide_3, "Relative Distance with Outlier Removal")
p4 <- generateplot(scag_noise_4, bignumbats_scags_wide_4, "Absolute Distance with Outlier Removal")

# Save plots
ggsave("figures/scagcomparison1.png", p1)
ggsave("figures/scagcomparison2.png", p2)
ggsave("figures/scagcomparison3.png", p3)
ggsave("figures/scagcomparison4.png", p4)
```

abs distance notes
- checked noise because if the scagnostics are being computed at stages, it would be better to have a smaller sample than the original data set
- It looks like sparse still has the sample size problem. Convex can technically identify the is also throwing weird values. Treating some as though the alphahull doesnt have an area
- Splines, and Dcor seem to be the best at identifying the shape
- Clumpy2 (i have no idea why) and monotonic are also doing an alright job
- Sparse can "see" the numbat, but it is somewhat unreliable because of its sample size dependency, outlying is similar but the reliability is questionable due to its large variance.
- check the alphahull area problem

relative distance notes
- seems like only dcor and sparse can see it... kind of?

# Tour with absolute value distnace
```{r}
#try with numbat data and lda 
datapartition <- createDataPartition(numbats_raw$group, p=0.2)$Resample1
numbat_subset <- numbats_raw[datapartition,]

# make matrix version of functions
tourr_clumpy <- function(mat){
  sc_clumpy(x= mat[,1], y=mat[,2])
}
tourr_splines <- function(mat){
  sc_splines(x= mat[,1], y=mat[,2])
}
tourr_convex <- function(mat){
  1- sc_convex(x= mat[,1], y=mat[,2])
}
tourr_dcor <- function(mat){
  sc_dcor(x= mat[,1], y=mat[,2])
}
tourr_sparse <- function(mat){
  sc_sparse(x= mat[,1], y=mat[,2])
}
tourr_monotonic <- function(mat){
  sc_monotonic(x= mat[,1], y=mat[,2])
}
tourr_outlying <- function(mat){
  1- sc_outlying(x= mat[,1], y=mat[,2])
}

# see if works with just numbat group
numbat_subset2 <- numbat_subset %>% filter(group=="A")

# does it see the numbat-ish?
animate_xy(numbat_subset2[, 1:10], guided_tour(tourr_splines)) #sometimes?
animate_xy(numbat_subset2[, 1:10], guided_tour(tourr_dcor), sphere = TRUE) # sometimes?
# animate_xy(numbat_subset2[, 1:10], guided_tour(tourr_clumpy), sphere = TRUE) #takes too long
animate_xy(numbat_subset2[, 1:10], guided_tour(tourr_convex), sphere = TRUE) #no
animate_xy(numbat_subset2[, 1:10], guided_tour(tourr_sparse), sphere = TRUE) #no
animate_xy(numbat_subset2[, 1:10], guided_tour(tourr_monotonic), sphere = TRUE) #kind of (splits clusters)
animate_xy(numbat_subset2[, 1:10], guided_tour(tourr_outlying), sphere = TRUE) #no
```

- This is basically a dead end
- Makes sense, the functions are being calculated on the entire thing rather than on the two groups, but im not sure it would find the numbat reguardless
- Even trying with just the data that contains the numbats
- When only given group A, splines can sepparate the two groups found in x4 most of the time. It also found the numbat shape on its second go, but that may have been a fluke because it didn't find it again.
- Clumpy can "see" the difference on the whole data set, but struggles a bit more on subsets. The measure seems to struggle a LOT with projection persuit, its esentially useless here lol.
- Convex can also "see" the difference on the whole data set, but it also has high variance on subsets. It does not succeed much here either.
- convex and clumpy are also very slow to calculate and the tour takes a while and makes my computer very hot.

# Tour with relative distances

```{r} 
#maxdist tourr:::
function (data, center = FALSE) 
{
    max(sqrt(rowSums(data^2)))
}
tourr:::max_dist


  
tourr_dcor <- function(cl){
  #this will only work on two groups
  classes <- unique(cl)
  #only work for 2, 
    # could do a for loop for more than 2 but not sure what to do 
    # with the splines values once i have them in that case. 
  function(mat){
    #each matrix group
    group1 <- mat[which(cl==classes[1]),]
    group2 <- mat[which(cl==classes[2]),]
  
    #scag value
    val1 <- sc_dcor(x= group1[,1], y= group1[,2])
    val2 <- sc_dcor(x= group2[,1], y= group2[,2])
  
    #final value (s1 - s2)/(s1 + s2)
    (val1 - val2)/(val1 + val2)
  }
}

#try tourr with this
animate_xy(numbat_subset[, 1:10], guided_tour(tourr_dcor(cl=numbat_subset$group)), col = numbat_subset$group)

```

- change distance with (s1 - s2)/(s1 + s2)
- this basically is just noise
- it isnt moving towards anthing.

# Try again after comparing distance metrics and outlier removal
- now using a global range (only relevant for scagnostic measures)
- also will try a different optimisation technique 

## Splines 
The measures biggest gap was with absolute difference, the relative difference was not really anything? (no wonder it struggled lol)
There was not much of a difference in outerlier removal
Also did a different optimisation method
```{r}
# first check one group
tourr_splines <- function(cl){
  classes <- unique(cl)
  function(mat){
    #each matrix group
    group1 <- mat[which(cl==classes[1]),]
    group2 <- mat[which(cl==classes[2]),]
  
    #scag value
    val1 <- sc_splines(x = group1[,1], y = group1[,2])
    val2 <- sc_splines(x = group2[,1], y = group2[,2])
  
    #final value (s1 - s2)/(s1 + s2)
    abs(val1-val2)
  }
}

animate_xy(numbat_subset[, 1:10], guided_tour(tourr_splines(cl=numbat_subset$group)), 
           sphere = TRUE, , col = numbat_subset$group)
```

## Clumpy2
```{r}
datapartition2 <- createDataPartition(numbats_raw$group, p=0.1)$Resample1
numbat_subset2 <- numbats_raw[datapartition2,]

# first check one group
tourr_clumpy <- function(cl){
  classes <- unique(cl)
  function(mat){
    #each matrix group
    group1 <- mat[which(cl==classes[1]),]
    group2 <- mat[which(cl==classes[2]),]
  
    #scag value
    val1 <- calc_scags(x= group1[,1], y= group1[,2], scags="clumpy2")$clumpy2
    val2 <- calc_scags(x= group2[,1], y= group2[,2], scags="clumpy2")$clumpy2
    
    #final val
    abs(val1-val2)
  }
}

animate_xy(numbat_subset2[, 1:10], guided_tour(tourr_clumpy(cl=numbat_subset2$group)), 
           sphere = TRUE, , col = numbat_subset2$group)
```


# Try to Visualise Optimisation
```{r, splines check}
#load ferrn library
library(ferrn)

#data
atapartition <- createDataPartition(numbats_raw$group, p=0.2)$Resample1
numbat_subset <- numbats_raw[datapartition,]
numbat_subset2 <- numbat_subset %>% filter(group=="A")

# matrix version of function
tourr_splines <- function(mat){
  sc_splines(x= mat[,1], y=mat[,2])
}

# does it see the numbat-ish?

animate(numbat_subset2[, 1:10],
        tour_path = guided_tour(tourr_splines),
        display = display_xy())

```

