---
title: "Try to Identify Numbat in Noise"
author: "Harriet Mason"
date: "07/12/2021"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load package and data
```{r}
# cassowaryr
library(cassowaryr)

#data wrangling
library(tidyverse)

# pairwise plots
library(GGally)
library(plotly)

#data partition 
library(caret)

# load data
numbats_raw <- read.csv("numbat.csv")

```

## Between Group Scagnostics
First calculate the scagnostics on the entire dataset (save data because this takes forever).
```{r, eval=FALSE}
# make group data frames
set.seed(1)
bignumbatsA <- filter(numbats_raw, group=="A") %>% select(-group)
bignumbatsB <- filter(numbats_raw, group=="B") %>% select(-group)

#calc_scags on both groups
bignumbatsA_scags <- calc_scags_wide(bignumbatsA)#, out.rm=FALSE)  
bignumbatsB_scags <- calc_scags_wide(bignumbatsB)#, out.rm=FALSE)

#combine into a single data frame
bignumbatsA_scags_long <- bignumbatsA_scags %>% 
  pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "A_value")

bignumbatsB_scags_long <- bignumbatsB_scags %>% 
  pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "B_value")

bignumbats_scags_long <- bignumbatsA_scags_long %>%
  left_join(bignumbatsB_scags_long, by=c("Var1", "Var2", "scag")) %>%
  mutate(group_diff = abs(A_value - B_value))

#make wide data
bignumbats_scags_wide <- bignumbats_scags_long %>%
  select(c(Var1, Var2, scag, group_diff)) %>%
  pivot_wider(id_cols=c(Var1, Var2), names_from=scag, values_from = group_diff)

#save data
saveRDS(bignumbats_scags_wide, "wholesetscags.rds")

```

```{r, eval=FALSE}
#look at SPLOM of scagnostics
splom_data <- bignumbats_scags_wide %>%
  mutate(lab = paste0(Var1, " , ", Var2)) %>%
  ungroup() %>%
  mutate(numbat=ifelse(lab == "x7 , x4", TRUE,FALSE)) %>%
  select(-c(Var1, Var2))

#interactive ggpairs
p <- ggpairs(splom_data, columns=c(1:11), aes(label=lab, colours=numbat)) +
  theme_minimal()
ggplotly(p) 
```

```{r}
# top pair for each scag
top_scags(bignumbats_scags_wide)
```

Scag Placings:
- Numbat 1st: Dcor, monotonic, outlying splines
- Numbat 2nd: Convex, Skinny, Sparse
- Numbat top 5: Skewed
- Numbat Nowhere: Euclid, Striated2, Stringy

Does it change with sample/sample size? Yes

Does it change with outlier removal? Yes
- x7 and x4 how top clumpy 2 difference???

Clumpy2 Specific
- identifies variable 7 as something interesting but not the combination. 
- Invites an investigation that looks at only one variable
- Alternatively could compare groups to the average across all variables? 
- Still maintains issue that one variable highly influences this scagnsotic, also 

Euclid
- Doesn't appear at all on euclidean distance
- also makes me think the clumpy value's sharp drop to 0 influences things like euclidean distance too much

## How much noise do we have on a 500 observation sample

```{r, eval=FALSE}
#subset variables
numbats_small <- numbats_raw %>%
  select(x1,x2,x4,x7, group)

# Set data
scag_data <- NULL

for(i in seq(20)){
  # load data
  set.seed(i)
  smallset <- createDataPartition(numbats_small$group, p=0.25)$Resample1
  numbatset <- numbats_small[smallset,]
  # make group data frames
  numbatsA <- filter(numbatset, group=="A") %>% select(-group)
  numbatsB <- filter(numbatset, group=="B") %>% select(-group)
  
  #calc_scags on both groups
  numbatsA_scags <- calc_scags_wide(numbatsA)  
  numbatsB_scags <- calc_scags_wide(numbatsB)
  
  #combine into a single data frame
  numbatsA_scags_long <- numbatsA_scags %>% 
    pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "A_value")
  
  numbatsB_scags_long <- numbatsB_scags %>% 
    pivot_longer(cols=-c(Var1, Var2), names_to = "scag", values_to = "B_value")
  
  numbats_scags_long <- numbatsA_scags_long %>%
    left_join(numbatsB_scags_long, by=c("Var1", "Var2", "scag")) %>%
    mutate(group_diff = abs(A_value - B_value))
  
  #make wide data
  numbats_scags_wide <- numbats_scags_long %>%
    select(c(Var1, Var2, scag, group_diff)) %>%
    pivot_wider(id_cols=c(Var1, Var2), names_from=scag, values_from = group_diff)
  
  # check 
  scag_data <- bind_rows(scag_data, numbats_scags_wide)
}

saveRDS(scag_data, "scagnostic_noise.rds")
```


```{r}
scag_data <- readRDS("scagnostic_noise.rds")
bignumbats_scags_wide <- readRDS("wholesetscags.rds")

#tidy true values
smallsetplots <- c("x2 , x1", "x4 , x1", "x4 , x2", 
                   "x7 , x1", "x7 , x2", "x7 , x4")
scag_true <- bignumbats_scags_wide %>% 
  mutate(lab = paste0(Var1, " , ", Var2)) %>%
  select(-c(Var1, Var2)) %>%
  filter(lab %in% smallsetplots)%>%
  pivot_longer(cols=-lab, names_to = "scag", values_to = "value")

#tidy data
scag_noise <- scag_data %>% 
  mutate(lab = paste0(Var1, " , ", Var2)) %>%
  select(-c(Var1, Var2)) %>%
  pivot_longer(cols=-lab, names_to = "scag", values_to = "value")

#plot
ggplot(scag_noise, aes(x=lab, y=value, colour=lab)) +
  geom_violin()+
  geom_point(alpha=0.5)+
  facet_wrap(~scag, scales = "free") +
  theme_minimal() +
  geom_point(data = scag_true, colour="black")
```

Notes
- checked noise because if the scagnostics are being computed at stages, it would be better to have a smaller sample than the original data set
- It looks like sparse still has the sample size problem. Convex can technically identify the is also throwing weird values. Treating some as though the alphahull doesnt have an area
- Splines, and Dcor seem to be the best at identifying the shape
- Clumpy2 (i have no idea why) and monotonic are also doing an alright job
- Sparse can "see" the numbat, but it is somewhat unreliable because of its sample size dependency, outlying is similar but the reliability is questionable due to its large variance.
- check the alphahull area problem
```{r}
plot(numbats_raw$x4, numbats_raw$x7)

# check convex value
#almost 0
calc_scags(bignumbatsA$x1, bignumbatsA$x2, scags="convex")
calc_scags(bignumbatsB$x1, bignumbatsB$x2, scags="convex")

#same as numbat
calc_scags(bignumbatsA$x7, bignumbatsA$x2, scags="convex")
calc_scags(bignumbatsB$x7, bignumbatsB$x2, scags="convex")
draw_alphahull(bignumbatsA$x2, bignumbatsA$x7)
draw_convexhull(bignumbatsA$x2, bignumbatsA$x7)
draw_alphahull(bignumbatsB$x4, bignumbatsB$x7)
draw_convexhull(bignumbatsB$x4, bignumbatsB$x7)

#numbat
calc_scags(bignumbatsA$x7, bignumbatsA$x4, scags="convex")
calc_scags(bignumbatsB$x7, bignumbatsB$x4, scags="convex")
draw_alphahull(bignumbatsA$x4, bignumbatsA$x7)
draw_convexhull(bignumbatsA$x4, bignumbatsA$x7)
draw_alphahull(bignumbatsB$x4, bignumbatsB$x7)
draw_convexhull(bignumbatsB$x4, bignumbatsB$x7)

#x1 and x2 have a 0 on "convex" but that shouldnt be true???

draw_alphahull(numbats_raw$x2, numbats_raw$x1, fill=TRUE)
# try with no outlier removal
draw_convexhull(numbats_raw$x2, numbats_raw$x1, fill=TRUE)
draw_alphahull(numbats_raw$x2, numbats_raw$x1, fill=TRUE, out.rm=FALSE)


```
- alpha hull in this plot has no fill. why is this happening?

- change distance with (s1 - s2)/(s1 + s2)

# Group vs Whole Data Scagnostics
```{r}

```
