---
title: "Question for Sherry"
author: "Harriet Mason"
date: "08/04/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r, loadlibraries}
library(tidyverse)
library(cassowaryr)
library(tourr)
library(ferrn)
```

# Issue
OK, so basically I can't use the scagnostic index function when I use the explore_space_pca() function. Since the scagnostic index is written up similar to lda_pp(), they are called similarly. I couldn't find an example for lda_pp() so I'm not sure if it is supposed to be called differently. 

# Data
 This is the data I'm working with and its trying to get to the projection that is just x1 and x4.
```{r simdata, fig.cap= "A scatter plot matrix of the simulated 'L-shape vs noise' data. The shape can be seen in the x1 vs x4 pairwise plot in group A."}
# set seed
set.seed(2020)

# get L-shape data
lshape <- features %>% 
  filter(feature=="l-shape")

# make tibble group a
feature_vs_noise_A <- tibble(x1 = 2.5* cassowaryr:::unitize(lshape$x),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(lshape$y),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("A",100)
                           )

# Make tibble group b
feature_vs_noise_B <- tibble(x1 = rnorm(100, 0.5, 1),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = rnorm(100, 0.5, 1),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("B",100)
                           )


# combine tibbles
feature_vs_noise <- bind_rows(feature_vs_noise_A, feature_vs_noise_B)

# final tibble and ggpairs check
p1 <- GGally::ggpairs(feature_vs_noise, columns = c(1:5),
                ggplot2::aes(colour = group, alpha=0.5))+ 
  theme_minimal()
  

p1
```

# Index Function
This is the index function I'm using. It basically just calculates the group differences in the scagnostic chosen for that projection. Takes a class argument like lda_pp(), but also has the scag option.

```{r, Function, echo=TRUE}
# Scag Projection Persuit index
scags_groups <- function(cl, scag){
  classes <- unique(cl)
  function(mat){
    #each matrix group
    group1 <- mat[which(cl==classes[1]),]
    group2 <- mat[which(cl==classes[2]),]
    indx <- list(
      "convex" = (sc_convex(x = group1[,1], y = group1[,2]) - sc_convex(x = group2[,1], y = group2[,2])),
      "outlying" = (sc_outlying(x = group1[,1], y = group1[,2]) - sc_outlying(x = group2[,1], y = group2[,2])),
      "splines" = (sc_splines(x = group1[,1], y = group1[,2]) - sc_splines(x = group2[,1], y = group2[,2])),
      "skinny" = (sc_skinny(x = group1[,1], y = group1[,2]) - sc_skinny(x = group2[,1], y = group2[,2])),
      "stringy" = (sc_stringy(x = group1[,1], y = group1[,2]) - sc_stringy(x = group2[,1], y = group2[,2])),
      "striated2" = (sc_striated2(x = group1[,1], y = group1[,2]) - sc_striated2(x = group2[,1], y = group2[,2])),
      "sparse" = (sc_sparse(x = group1[,1], y = group1[,2]) - sc_sparse(x = group2[,1], y = group2[,2])),
      "skewed" = (sc_skewed(x = group1[,1], y = group1[,2]) - sc_skewed(x = group2[,1], y = group2[,2])),
      "splines" = (sc_splines(x = group1[,1], y = group1[,2]) - sc_splines(x = group2[,1], y = group2[,2]))
      )
    
    #final value
    abs(indx[[scag]])
    
  }
}
```

It works fine if I run it as a tour projection persuit index, and occasionally finds the shape we are looking for (so its all good in that reguard).

```{r}
set.seed(7)
a <- animate_xy(feature_vs_noise[,1:5],
                tour_path = guided_tour(scags_groups(feature_vs_noise$group, "convex"),
                                        search_f = search_better,
                                        alpha=0.75),
                col = feature_vs_noise$group)
set.seed(6)
b <- animate_xy(feature_vs_noise[,1:5],
                tour_path = guided_tour(scags_groups(feature_vs_noise$group, "convex"),
                                        search_f = search_better,
                                        alpha=0.75),
                col = feature_vs_noise$group)
```

This funtion just shows the best projection of the tours above so I could check if it found the right shape. 

```{r}

last_frame <- function(numeric_data, group_data, anim_data){
  # make data into matrix
  mat_data <- as.matrix(numeric_data)
  mat_basis <- matrix(unlist(get_best(anim_data)$basis), ncol=2, nrow=5)
  
  # transform data to basis
  plot_data <- as_tibble(mat_data %*% mat_basis) %>%
    rename(x = V1, y = V2)
  plot_data$group <- group_data
  
  # get basis data for segment plots
  basis_data <- as_tibble(matrix(unlist(get_best(anim_data)$basis), ncol=2, nrow=5)) %>%
    rename(xend = V1, yend = V2) %>%
    mutate(x=0, y=0, xend=5*xend, yend=5*yend)
  basis_data$variable = c("x1", "x2", "x3", "x4", "x5")
  
  #plot ggplot
  ggplot(plot_data, aes(x,y)) + 
    geom_point(aes(colour=group)) +
    theme_classic() +
    geom_segment(data=basis_data, aes(x=x, y=y, xend=xend, yend=yend)) +
    geom_text(data=basis_data, aes(x=xend, y=yend, label=variable)) +
    theme(axis.line = element_blank(),
          axis.ticks = element_blank(),
          axis.text = element_blank())

}
last_frame(feature_vs_noise[1:5], feature_vs_noise$group, a)
last_frame(feature_vs_noise[1:5], feature_vs_noise$group, b)

```

The trace plot functions from the ferrn package work, so the problem isn't the a or b data sets.

```{r}
a %>% 
  explore_trace_interp() + 
  scale_color_continuous_botanical() + ggtitle("Seed 7")

b %>% 
  explore_trace_interp() + 
  scale_color_continuous_botanical() + ggtitle("Seed 6")
```

But the explore_space_pca() function throws an error "group1[, 1] : incorrect number of dimensions". I know that this error is from is code from inside the scagnsotic index function when it splits the data based upon the groups. I can't remove the groups or scag function from the index, so I'm not sure how to use it.

```{r}
bind_rows(a, b) %>%
  bind_theoretical(matrix(c(1, 0, 0, 1, 0), nrow = 5),
                   index = scags_groups(feature_vs_noise$group, "convex"), 
                   raw_data = feature_vs_noise[, 1:5]) %>% 
  explore_space_pca(group = method, details = TRUE)  +
  scale_color_discrete_botanical()
```

The example works fine so I dont think the problem is my computer running ferrn haha.

```{r}
# Example
dplyr::bind_rows(holes_1d_geo, holes_1d_better) %>%
  bind_theoretical(matrix(c(0, 1, 0, 0, 0), nrow = 5),
    index = tourr::holes(), raw_data = boa5
  ) %>%
  explore_space_pca(group = method, details = TRUE) +
  scale_color_discrete_botanical()
```
