---
title: "Projection Pursuit Write Up"
author: "Harriet Mason"
date: "21/04/2022"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, loadlibraries}
library(tidyverse)
library(cassowaryr)
library(tourr)
library(ferrn)
library(stringr)
```

# The feature vs noise data

We want to identify if scagnostics can be used in projection persuit to find interesting shapes in linear projections of a data set. In order to test this, we created a data set with two groups where one pair of variables has a "shape" in one of the groups, and all other variables are normally distributed. Here the scagnsotics will have to find the L-shape in group A in the x1 and x4 pairwise plot that can be seen in the scatter plot matrix (SPLOM) below. If the scagnsotics can be used in projection persuit, they should be able to identify which group and variable combination shows the abonormal shape.

```{r simdata, fig.cap= "A scatter plot matrix of the simulated 'L-shape vs noise' data. The shape can be seen in the x1 vs x4 pairwise plot in group A."}
# set seed
set.seed(2020)

# get L-shape data
lshape <- features %>% 
  filter(feature=="l-shape")

# make tibble group a
feature_vs_noise_A <- tibble(x1 = 2.5* cassowaryr:::unitize(lshape$x),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = 2.5* cassowaryr:::unitize(lshape$y),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("A",100)
                           )

# Make tibble group b
feature_vs_noise_B <- tibble(x1 = rnorm(100, 0.5, 1),
                           x2 = rnorm(100, 0.5, 1),
                           x3 = rnorm(100, 0.5, 1),
                           x4 = rnorm(100, 0.5, 1),
                           x5 = rnorm(100, 0.5, 1),
                           group = rep("B",100)
                           )


# combine tibbles
feature_vs_noise <- bind_rows(feature_vs_noise_A, feature_vs_noise_B)

# final tibble and ggpairs check
p1 <- GGally::ggpairs(feature_vs_noise, columns = c(1:5),
                ggplot2::aes(colour = group, alpha=0.5))+ 
  theme_minimal()
  

p1
```

Any scagnsotic that can identify this shape in projection pursuit should also be able to pick the shape from a collection pairwise plots. This is how we are going to select the projection pursuit index. Here we calculate each scagnostic, for each possible pairwise plot, separately for each group. We then calculate the absolute difference between these values for each group. The scagnostics that have their maximum group difference on the x1 and x4 pair can "see" the shape. The table below has the difference between the groups and the associated variable pair for each scagnostic. The scagnostics convex and skinny are the best choice for projection pursuit and we will continue with those measures.

```{r, eval=FALSE scagcals, fig.cap= "The top pairs of variables. Convex and skinny do the best job of seeing the L-shape in group A. }
# calculate scagnostics
scags_feature_vs_noise_A <- calc_scags_wide(feature_vs_noise_A[,1:5]) %>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="A_value")

scags_feature_vs_noise_B <- calc_scags_wide(feature_vs_noise_B[,1:5])%>%
  pivot_longer(cols=3:13, names_to = "scag", values_to="B_value")

# Combine Data
scags_feature_vs_noise_dif <- scags_feature_vs_noise_A %>%
  left_join(scags_feature_vs_noise_B) %>%
  mutate(dif_value = abs(A_value - B_value)) 

scag_df <- scags_feature_vs_noise_dif %>%
  select(-c(A_value, B_value)) %>%
  pivot_wider(id_cols = c(Var1, Var2), names_from=scag, values_from=dif_value)
  
top_scags(scag_df) %>%
  arrange(desc(value))
  
```

## 2) The Scagnostics Index Function 

Scagnostics can be easily integrated into the `tourr` projection pursuit by defining an index to be optimised in the guided tour. The scagnostic index function takes a vector containing the class of each observation (cl) and a string depicting which scagnostic should be used (scag). The index maximises the absolute difference in that scagnostic across the groups, the same calculation we used to identify which scagnsotic to use in projection pursuit. Only the scagnsotics that are rationally invariant and computationally efficient enough to be used in projection pursuit are options in this function. 

```{r, Function, echo=TRUE}
# Scag Projection Persuit index
scags_groups <- function(cl, scag){
  classes <- unique(cl)
  function(mat){
    #each matrix group
    group1 <- mat[which(cl==classes[1]),]
    group2 <- mat[which(cl==classes[2]),]
    indx <- list(
      "convex" = (sc_convex(x = group1[,1], y = group1[,2]) - sc_convex(x = group2[,1], y = group2[,2])),
      "outlying" = (sc_outlying(x = group1[,1], y = group1[,2]) - sc_outlying(x = group2[,1], y = group2[,2])),
      "splines" = (sc_splines(x = group1[,1], y = group1[,2]) - sc_splines(x = group2[,1], y = group2[,2])),
      "skinny" = (sc_skinny(x = group1[,1], y = group1[,2]) - sc_skinny(x = group2[,1], y = group2[,2])),
      "stringy" = (sc_stringy(x = group1[,1], y = group1[,2]) - sc_stringy(x = group2[,1], y = group2[,2])),
      "striated2" = (sc_striated2(x = group1[,1], y = group1[,2]) - sc_striated2(x = group2[,1], y = group2[,2])),
      "sparse" = (sc_sparse(x = group1[,1], y = group1[,2]) - sc_sparse(x = group2[,1], y = group2[,2])),
      "skewed" = (sc_skewed(x = group1[,1], y = group1[,2]) - sc_skewed(x = group2[,1], y = group2[,2])),
      "splines" = (sc_splines(x = group1[,1], y = group1[,2]) - sc_splines(x = group2[,1], y = group2[,2]))
      )
    
    #final value
    abs(indx[[scag]])
    
  }
}
```

## 3) Tuning the Projection Persuit Indexes
The scagnostics can find the correct shape more often than not, but this is only with some adjustments to the optimization parameters. Below is an example of the tour projection persuit using the skinny scagnostic to find the l-shape projection using adjusted parameters, and failling to find the l-shape in the data with the default parameters.

```{r, Tourgrid}
# First make a function that animates the tour. This can be done in a for loop but for the large calculations im doing this makes it easier

# Animate Tour Function: Function just runs a tour and saves it as as an r object
anim_tourr <- function(numeric_data, group_vector, scag, search, seed, alpha, max.tries, filename){
  #set seed
  set.seed(seed)
  
  # perform search by calling function from list
  srch <- list(
    "bet" =  animate_xy(numeric_data,
                  tour_path = guided_tour(scags_groups(group_vector, scag),
                                          search_f = search_better,
                                          alpha = alpha,
                                          max.tries=max.tries),
                  col = group_vector),
    "bet_rand" = animate_xy(numeric_data,
                  tour_path = guided_tour(scags_groups(group_vector, scag),
                                          search_f = search_better_random,
                                          alpha = alpha,
                                          max.tries=max.tries),
                  col = group_vector)
    )
  
  #save object and file name
  a <- srch[[search]]
  #filename <- paste(filestart, scag, search, alpha, max.tries, seed, ".rds", sep="_")
  
  # add in identifiers for explort_pca plot
  a <- a %>%
    mutate(scag = scag,
           origalpha = alpha,
           max.tries = max.tries,
           seed = seed
           )
  # save r object
  saveRDS(a, filename)
}

# set up a function that lets us run that animate tour function over a grid of values
grid_tourr <- function(numeric_data, group_vector, scags, searches, seeds, alpha, max.tries, filestart, ...){
  #make expanded tibble so that we can search a grid of values
  expand.grid(scags, searches, seeds, alpha, max.tries, stringsAsFactors=FALSE) %>%
    # add file names for saving
    mutate(filename = paste(filestart, Var1, Var2, Var4, Var5, Var3, ".rds", sep="_")) %>%
    # group by and summarise to run save on every value
    group_by(Var1, Var2, Var4, Var5, Var3) %>%
    summarise(anim_tourr(numeric_data, group_vector, scag = Var1,  search = Var2, seed = Var3, 
                         alpha = Var4, max.tries=Var5, filename = filename))
}

# Make a small grid to run all the tours on
scags <- c("skinny")
max.tries <- 100
alpha <- c(0.1, 0.7)
seeds <- 1

# Run Tours
grid_tourr(feature_vs_noise[,1:5], feature_vs_noise$group, scags, "bet", seeds, alpha, max.tries, "writeuptours/")


# read them back into R and combine into one tibble


# Explore PCA plot

```

```{r, eval=FALSE}
# GEODESIC WORK
set.seed(2)
m <- animate_xy(feature_vs_noise[1:5],
                tour_path = guided_tour(scags_groups(feature_vs_noise$group, "convex")),
                delta=0.05,
                col = feature_vs_noise$group)
 # BETTER WORK
set.seed(7)
m <- animate_xy(feature_vs_noise[1:5],
                tour_path = guided_tour(scags_groups(feature_vs_noise$group, "convex"),
                                        search_f = search_better,
                                        alpha=0.75),
                col = feature_vs_noise$group)

# DEFAULT DOES NOT WORK
animate_xy(feature_vs_noise[1:5],
           tour_path = guided_tour(scags_groups(feature_vs_noise$group, "convex"),
                                   search_f = search_better),
           col = feature_vs_noise$group)
```

## 4) Trace plots and Results
## 4) Improvements / Which scagnostics can be used in projection persuit + why
- we have worked backwards here to show that the scagnsotics can be used to identify shapes and tinker the optimiser. When using scagnostics in projection persuit you would know what kind of shape you are looking for, and do a projection persuit to see if it exists anywhere in the data (While the "correct: projection would not usually be known to us, since we are trying to see if the measures work at all, we will be working backwards from the final results)
- clumpy2 is too computationally expensive
- not scaled globally 0 to 1
- scagnstics do not actually scale 0 to 1, therefor the scagnostic with the biggest difference may not be convex, but sparse. cannot tell because of scaling issue.
- more variables
- different shapes
